### 2019-12-08 presentations/2019-12-09/presentation.tex
Preparing presentation for tomorrow 1

\documentclass[10pt]{beamer}

%packages
\usepackage{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,tabu}
\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{matrix,chains,positioning,decorations.pathreplacing,arrows}
\usetikzlibrary{fit,shapes}
\usetikzlibrary{calc,shadings}
\usepackage{pgfplots}
\usepackage{colortbl}
\usepackage{eurosym}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{tabularx}

%definitions
\usepackage{algorithm,algorithmic}

%theme
\usetheme{Dresden}
\usecolortheme{rose}
\useoutertheme{tree}

%environments
\newenvironment{ExampleGer}{\begin{exampleblock}{Beispiel}}{\end{exampleblock}}

\newenvironment{customlegend}[1][]{%
	\begingroup
	% inits/clears the lists (which might be populated from previous
	% axes):
	\csname pgfplots@init@cleared@structures\endcsname
	\pgfplotsset{#1}%
}{%
	% draws the legend:
	\csname pgfplots@createlegend\endcsname
	\endgroup
}%

%definitions
\def\addlegendimage{\csname pgfplots@addlegendimage\endcsname}
% definition to insert numbers
\pgfkeys{/pgfplots/number in legend/.style={%
		/pgfplots/legend image code/.code={%
			\node at (0.295,-0.0225){#1};
		},%
	},
}

%general
\title{Combining bottom up and top down approaches for Drug-Target-Interaction prediction}
\author{Tilman Hinnerichs}
\institute{BORG - KAUST}
\date{December 09, 2019}

%presentation
\expandafter\def\expandafter\insertshorttitle\expandafter{%
	\insertshorttitle\hspace{5cm}%
	\insertframenumber\,/\,\inserttotalframenumber
}

\definecolor{myblue}{RGB}{80,80,160}
\definecolor{mygreen}{RGB}{80,160,80}
\definecolor{myturk}{RGB}{160,160,160}

\begin{document}

\begin{frame}
	\titlepage
\end{frame}

\begin{frame}{Outline}
	\setbeamertemplate{section in toc}[sections numbered]
	\tableofcontents
\end{frame}

\section{Problem Description}
\begin{frame}{Problem Description}
	Prediction over bipartite graph:

	\begin{tikzpicture}[thick,
	every node/.style={draw,circle},
	fsnode/.style={fill=myblue},
	fs1node/.style={fill=myturk},
	ssnode/.style={fill=mygreen},
	%every fit/.style={ellipse,draw,inner sep=-1pt,text width=2cm},
	%->,shorten >= 3pt,shorten <= 3pt
	]

	% the vertices of U
	\begin{scope}[start chain=going below,node distance=5mm]
	\foreach \i in {1,2,...,3}
	\node[fsnode,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of U1
	\begin{scope}[node distance=5mm]
	\foreach \i in {4,5}
	\node[fs1node,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of V
	\begin{scope}[xshift=4cm,yshift=-0.25cm,start chain=going below,node distance=7mm]
	\foreach \i in {1,2,...,4}
	\node[ssnode,on chain] (s\i) [label=right: T\i] {};
	\end{scope}

	% the set U
	%\node [myblue, fit=(f1) (f3)] {};
	\node [white,fit=(f1) (f5),label=above:$Drugs$] {};

	% the set V
	\node [white,fit=(s1) (s4),label=above:$Proteins$] {};

	% the edges
	\draw (f1) -- (s1);
	\draw (s1) -- (f2);
	\draw (f2) -- (s2);
	\draw (s2) -- (f3);
	\draw (s3) -- (f3);
	\draw (f3) -- (s4);
	\draw (s4) -- (f5);
	\draw (f5) -- (s1);
	\end{tikzpicture}

\end{frame}

\section{Recent approaches}
\begin{frame}{Classification of recent approaches\footnote{Chen Wang et al., Briefings in Bioinformatics, 2018}\footnote{Yu Ding, Briefings in Bioinformatics, 2019}}
	\begin{table}
		\begin{tabularx}{\textwidth}{|>{\setlength\hsize{.5\hsize}\setlength\linewidth{\hsize}}X|>{\setlength\hsize{1.25\hsize}\setlength\linewidth{\hsize}}X|>{\setlength\hsize{1.25\hsize}\setlength\linewidth{\hsize}}X|}
			\hline
			&Drugs&Protein\\
			\hline
			bottom-up&
			\begin{itemize}
				\item GCN over molecules
				\item drug similarity
			\end{itemize}&
			\begin{itemize}
				\item secondary structure prediction
				\item contact prediction
				\item convolution over amino acid sequences
			\end{itemize}\\
			\hline
			top-down&
			\begin{itemize}
				\item network approaches
				\item drug similarity
			\end{itemize}&
			\begin{itemize}
				\item protein similarity
			\end{itemize}\\
			\hline
		\end{tabularx}
	\end{table}	
\end{frame}

\begin{frame}{Problems of recent approaches}
	Main issues:
	\begin{itemize}
		\item lack of generalization
		\item Usually only top-down or bottom-up
	\end{itemize}
	\pause
	\begin{tikzpicture}[thick,
	every node/.style={draw,circle},
	fsnode/.style={fill=myblue},
	fs1node/.style={fill=myturk},
	ssnode/.style={fill=mygreen},
	%every fit/.style={ellipse,draw,inner sep=-1pt,text width=2cm},
	%->,shorten >= 3pt,shorten <= 3pt
	]

	% the vertices of U
	\begin{scope}[start chain=going below,node distance=5mm]
	\foreach \i in {1,2,...,3}
	\node[fsnode,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of U1
	\begin{scope}[node distance=5mm]
	\foreach \i in {4,5}
	\node[fs1node,on chain] (f\i) [label=left: D\i] {};
	\end{scope}

	% the vertices of V
	\begin{scope}[xshift=4cm,yshift=-0.25cm,start chain=going below,node distance=7mm]
	\foreach \i in {1,2,...,4}
	\node[ssnode,on chain] (s\i) [label=right: T\i] {};
	\end{scope}

	% the set U
	%\node [myblue, fit=(f1) (f3)] {};
	\node [white,fit=(f1) (f5),label=above:$Drugs$] {};

	% the set V
	\node [white,fit=(s1) (s4),label=above:$Proteins$] {};

	% the edges
	\draw (f1) -- (s1);
	\draw (s1) -- (f2);
	\draw (f2) -- (s2);
	\draw (s2) -- (f3);
	\draw (s3) -- (f3);
	\draw (f3) -- (s4);
	\draw (s4) -- (f5);
	\draw (f5) -- (s1);
	\end{tikzpicture}

\end{frame}

\section{A medium approach to DTI prediction}

\begin{frame}{Full citations}
	\footnotesize
	\begin{itemize}
		\item Yu Ding, Hong Wang, Hewei Zheng, Lianzong Wang, Guosi Zhang, Jiaxin Yang, Xiaoyan Lu, Yu Bai, Haotian Zhang, Jing Li, Wenyan Gao, Fukun Chen, Shui Hu, Jingqi Wu, Liangde Xu, Evaluation of drug efficacy based on the spatial position comparison of drug–target interaction centers, Briefings in Bioinformatics, , bbz024, https://doi.org/10.1093/bib/bbz024
		\item Chen Wang, Lukasz Kurgan, Review and comparative assessment of similarity-based methods for prediction of drug–protein interactions in the druggable human proteome, Briefings in Bioinformatics, , bby069, https://doi.org/10.1093/bib/bby069
	\end{itemize}
\end{frame}

\end{document}