### 2021-04-15 paper/main.tex
Finished abstract, updated README

\documentclass{bioinfo}
\usepackage[english]{babel}
\copyrightyear{2015} \pubyear{2015}
\usepackage{natbib}
\bibliographystyle{natbib}
\usepackage{url}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\renewcommand{\cite}{\citep}
\usepackage{todonotes}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}
\newcommand{\name}{DTI-Voodoo}

\title[\name{}]{\name{}: Combining Bottom-up and top-down approaches
  through graph learning over interaction networks for
  drug-target-interaction prediction} \author[Sample
\textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert
  Hoehndorf\,$^{\text{\sfb 2}}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
  $^{\text{\sf 2}}$Computational Bioscience Research Center, Computer,
  Electrical and Mathematical Sciences \& Engineering Division, King
  Abdullah University of Science and Technology, 4700 King Abdullah
  University of Science and Technology, Thuwal 23955, Saudi Arabia.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} \textit{In silico} drug--target interaction (DTI) prediction is crucial for drug discovery and down-stream tasks, such as drug indication and drug--disease prediction. However, DTI datasets are prone to certain biases, skewing the effective predictive expressiveness over those datasets. Additionally, only few approaches combine phenotypical and molecular features for DTI prediction.\\
\textbf{Results:} We present \name{} to combine molecular features and functional information with protein--protein interaction networks under usage of graph convolutional neural networks, showing that phenotypic information localizes on the interaction network. We hereby focus on protein-centric evaluation, predicting drugs that may target specific targets. We further show that common drug--target interaction datasets contain intrinsic biases, majorly affecting predictive performance of compared models. Additionally, we propose an modified evaluation scheme and metric to circumvent such skews and to better reflect the problem DTI prediction methods aim to solve. \\
\textbf{Availability:} Source code and data used in \name{} are available at \url{https://github.com/THinnerichs/DTI-VOODOO}\\
\textbf{Contact:} \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}\\
\textbf{Supplementary information:}10264703 Supplementary data are available at \textit{Bioinformatics}
online.}

\maketitle
\section{Introduction}

% In history, traditional remedies, that were known for their medicinal
% properties lead to drugs by extraction of the functional
% ingredients. Alternatively, characteristics and features of potential
% drugs were detected by accident like in the case of penicillin. More
% recently, biological drug targets can be found \textit{in silico}
% through discovery of suitable computational predictors.

% Knowledge about those links between compounds and their
% target proteins help in an array of medical and pharmaceutical
% studies. Additionally, those associations can be utilized to identify
% disease specific targets, leading to desirable therapeutic effects.

% With the rapidly growing field of machine learning approaches and
% their application to bioscientifical problems in the realm of
% bioinformatics, different kinds of data, such as long DNA sequences
% could be utilized for feature generation, while rapid advances were
% made. 
Identifying drug--target interactions (DTIs) is important in drug
repurposing and repositioning, and in the exploration of novel drug
candidates and their interaction partners. Computational methods are
widely applied to predict DTIs and many computational methods have
been developed.  These methods can be broadly classified into
``top-down'' and ``bottom-up'' approaches. Top-down approaches start
form observable characteristics resulting from a drug--target
interaction, such as side-effects or the diseases treated by a drug,
and infer likely molecular mechanisms (i.e., the interaction) using
these observations.
% and indications represented by knowledge graphs or ontologies, induced by a drug and infer targets based on the likely molecular mechanisms that result in these phenotypes
Bottom-up approaches start from molecular features such as molecular
structure or fingerprints associated with drug and protein, and
predict interactions from this information.

Both bottom-up and top-down approaches to drug--target interaction
prediction have some advantages and limitations. Generally, bottom-up
methods face the challenge to predict whether a structure binds to a
protein given their molecular properties; whether two entities
interact depends not only on the molecular structure of the entities
(where binding sites and molecular forces need to be determined for
accurate prediction) but also properties such as where a protein is
expressed or in which celltypes and anatomical structures. Top-down
methods use information for DTI prediction such as side-effect
similarity \cite{Campillos2008} that is largely complementary to
knowledge gained from molecular properties. While methods tha3t rely on
molecular information are directly predicting whether two molecular
could interact, top-down methods rely on more indirect means and 
infer a DTI from observable effects resulting from the interaction.

Both approaches may be used in conjunction with network inference
\cite{}. Biological networks used for DTI prediction include
protein--protein interaction networks \cite{} and networks including
several other types of biological relations, including similarity
between represented entities \cite{Gottlieb2011}. Network-based DTI
prediction methods use the guilt-by-association principle
\cite{Oliver2000}
and assume that a protein is a likely target for a drug if many of the
protein's neighbors in the interaction network are targets of the drug
\cite{Gillis2012}. Network-based methods have been applied successfully
to DTI prediction. However, if DTIs are taken as direct physical
interactions between a drug and protein, it remains an unresolved
question whether the network-based guilt-by-association hypothesis is
true, or whether an interaction of a drug and protein dys-regulates
several of the protein's interaction partners, and therefore resulting
in downstream effects that are not direct interactions but only
downstream consequences of an interaction.

Progress in machine learning using graph neural networks can allow us
to test this hypothesis and combine both bottom-up and top-down
features with a network in a single machine learning model. 
In particular, Graph Convolutional Networks \cite{GCNConv} and
their variants using different types of kernels \cite{ChebConv,
  ARMAConv}, including attention mechanisms \cite{GATConv}, and
different forms of exploring node neighborhoods \cite{APPNPConv,
  SAGEConv} can combine different types of features and graph-based
information. % While based on diverse systems,
% they can be relevant for testing distinct hypothesis for given
% graphs. While convolutional filters are suitable for finding patterns
% among the given graph, attention mechanisms are more relevant for
% discovery of important regions within. Lately, graph learning
% approaches found application for computing compound representations
% for DTI prediction.
They have previously been applied for a number of tasks, including
prediction of protein functions \cite{Zitnik2017}, cancer drug response \cite{Liu2020} and drug--target affinity prediction \cite{GraphDTA2020}.

One challenge in DTI prediction are potential biases resulting from
the underlying datasets \citep{Pahikkala2014} which may affect model
evaluation and comparison. First, novel drugs are often designed by
altering non-functional components of a drug, leading to two and more
very similar drugs designed to target the same proteins
\cite{Overington2006}. This can result in a bias when it leads to
hidden duplicates or highly similar compounds that are distributed
among training and evaluation dataset, resulting in a better
(measured) predictive performance than would be expected when the
model is applied to identify drugs that target a protein for which no
drugs yet exist. Second, some proteins (which we call \textit{hub
  proteins}) have significantly more known interactions with drugs
than others. In the STITCH database, $5\%$ of the proteins have $40\%$
of the interactions, and similar distributions are present in other
datasets \cite{Drugbank2007, Drugbank2017}; preferentially predicting
these proteins may increase predictive performance while not
reflecting the actual performance when applied to a new protein (i.e.,
a protein for which no interactions are known). These differences in
the number of drugs targeting certain proteins may be the result of
study bias where more ``valuable'' proteins have more drugs designed
to target them due to their involvement in more common diseases (or
diseases for which drugs can be more profitably marketed). This may
affect common evaluation schemes where it is possible
to exploit these biases within DTI prediction \cite{Survey2018}.

We developed \name{} as a method for predicting DTIs. We use an
ontology-based machine learning method \cite{DL2vec2020} to encode phenotypic
consequences of DTIs and deep learning methods to encode molecular
features. We combine both using a protein interaction network which we
exploit using a graph neural network. We use this model to test
whether molecular or phenotype features benefit from the network
information and find that only phenotype features localize on the
graph whereas molecular features do not. We further evaluate and
compare \name{} against several DTI prediction methods and
demonstrate a substantial improvement of \name{} over the state of the
art in predicting drugs that target a protein. We also identify and
characterize several biases in both training and evaluating DTI
prediction methods, and make recommendations on how to avoid
them. \name{} is available as Free Software at \url{https://github.com/THinnerichs/DTI-VOODOO}.

% In order to design representations concerning both top-down and bottom-up features for proteins and drugs,respectively, we utilize protein--protein interaction (PPI) networks which have shown great results in protein function prediction \cite{Vazquez2003} and antiviral drug--target discovery \cite{Ackerman2019}. 
% However, only few applied this context to the task of DTI prediction in humans.

%\todo[inline]{Somehow smooth this transition}

\enlargethispage{12pt}

\section{Methods}
\subsection{Problem Description}
\name{} aims to solve the following problem: for a given drug and a
given protein we want to determine whether those interact or not.  We
do not differentiate between types of interaction such as activation
and inhibition, and do not predict the strength of the interaction.
We treat all drug--protein pairs without a known interaction as
negatives and therefore formulate the problem as a binary
classification task.

\subsection{Datasets}
We obtain a dataset consisting of 12,884 human proteins with 
340,627 links from STRING \citep{STRINGv10}. For the drug-target
interactions, we use 229,870 links from the STITCH database
\citep{STITCHv5}. As both STRING and STITCH provide confidence scores
for each association, we filtered them as advised by a threshold of
$700$, therefore retaining only high-confidence interactions.

We utilize the PhenomeNET ontology \citep{PhenomeNET2011}, an ontology
integrating ontologies such as the Human Phenotype Ontology
\citep{HPO2018}, Gene Ontology \cite{GOoriginal2000, GOrecent2020},
Mammalian Phenotype Ontology \citep{MP2009} and several others.  We
obtained side effects and their links to drugs from SIDER
\citep{SIDER}; SIDER contains side effects encoded using identifiers
from the MedDRA database \citep{MedDRA}. We mapped side effects to the
PhenomeNET ontology using the \textit{Phenomebrowser.net}, which
provides a SPARQL query endpoint for the mentioned resources.

We only use proteins in our analysis that have at least one link in
either STITCH or STRING, and drugs with at least one side effect and
one existing target. Therefore, the intersection between these
resources yields 1,428 drugs and 7,368 human proteins with 32,212
interactions for the training phase. We provide links to and methods
for obtaining and processing the necessary data on Github.

For comparative evaluation, we use the gold standard dataset
introduced by \citet{Yamanishi2008} consisting of 1,923 interactions
between 708 drugs and 1,512 proteins, and the BioSnap dataset
\cite{BioSnap2018} which consists of 5,017 drug nodes, 2,324 gene
nodes and 15,138 edges.

\subsection{Model}

Our model combines ``top-down'' and ``bottom-up'' information for
drug--target identification. We consider an approach to be
``top-down'' when observable characteristics of either a drug (such as
a drug effect) or protein (such as a protein function, or phenotypes
resulting from a loss of function) are used to provide information
about a molecular mechanisms; we consider an approach ``bottom-up''
when structural or other molecular information is used to determine a
mechanism.  In order to build a method that incorporates both top-down
and bottom-up features, we first create a model for each type of
feature separately.  As features for the bottom-up model, we use
features derived from molecular structures of drugs from the
\textit{SmilesTransformer} \citep{SmilesTransformer} and molecular
features for proteins from \textit{DeepGOPlus}
\citep{DeepGoPlus}. \textit{SmilesTransformer} is an autoencoder
trained over the SMILES strings, and therefore captures (some aspects
of) the molecular organization of each drug in an unsupervised manner.
\textit{DeepGOPlus} provides features derived from protein amino acid
sequences which are useful to predict protein function.
% Thus, both embeddings seem to suitably supplement
% the following ontology based representations.

As phenotypes and functions are encoded through ontologies, we use
DL2Vec \citep{DL2vec2020} to obtain ontology based representations for
use as top-down features. DL2vec constructs a graph by introducing
nodes for each ontology class and edges for ontology axioms, followed
by random walks starting from each node in the graph. These walks are
encoded using a Word2vec \citep{Word2vec2013} model. Therefore, DL2Vec
generates representations that can encode drug effects or protein
functions while preserving their semantic neighbourhood within that
graph.

% The overall structure of the ontology can be seen in
% Figure~\ref{fig:Onto}.

% \todo[inline]{Consider supplement, or combine with other figures.}
% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=0.5\textwidth]{figures/drug_protein_ontology_network.png}}
% 	\caption{Drugs and proteins with annotations to SiDER and PhenomeNET}
% 	\label{fig:Onto}
% \end{figure}

\subsubsection{Half-twin neural networks and feature transformation}

As we want to learn from the similarity of drug side effects and
protein phenotypes, we use a deep half-twin neural network with a
contrastive loss using cosine similarity.  A half-twin neural network
aims to learn a similarity between two embeddings of the same
dimension.  The precomputed embeddings may have different
dimensionality and we first process them using a fully connected
neural network layer which takes as input an embedding and outputs a
representation of a particular size, i.e., we use this layer as a
trainable feature transformation and apply it to reduce the
representation size of the embeddings for drugs and proteins
separately.  An example structure for both types of features can be
found in Figure~\ref{fig:HalfTwinNetwork}.
% While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
The use of this trainable feature transformation layer enables
flexible experimentation as both ontology and molecular feature for
both drugs and proteins are reduced to the same dimensionality for
varying sizes of inputs; this allows for a high amount of modularity
across different experimental setups by adding different kinds of
features into the model. Additionally, the generated features may be
used for other tasks. % ; for
% example, the ontology LFT can be reused for a variety of DL2vec-based
% features with respect to other ontologies and hypotheses
We follow the results of \textit{DL2vec} \cite{} and use as
activation function $\sigma := \mathrm{LeakyReLU}$ which leads to
improved performance compared to other activation functions.

% \todo[inline]{Combine in some larger figure}
% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=0.35\textwidth]{figures/siamese_network.png}}
% 	\caption{Half-twin network applied to molecular and DL2vec
%           features, utilizing deep learnable feature transformations
%           (LFT). The similarity function $\otimes$ yields the
%           similarity between both transformed embeddings e.g. by
%           computing the cosine similarity.}
% 	\label{fig:HalfTwinNetwork}
% \end{figure}

\subsubsection{Graph convolutional layers}

We include these molecular and ontology-based sub-models within a
graph neural network (GNN) \cite{}. The graph underlying the GNN is
based on the protein--protein interaction (PPI) graph. The PPI dataset
is represented by a graph $G=(V,E)$, where each protein is represented
by a vertex $v\in V$, and each edge $e\in E\subseteq V\times V$
represents an interaction between two proteins. Additionally, we
introduce a mapping $x:V\rightarrow\mathbb{R}^{d}$ projecting each
vertex $v$ to its node feature $x_v := x(v)$, where $d$ denotes the
dimensionality of the node features.

% As described before, graph convolution has shown significant
% performance increase in a variety of tasks. While there are various
% methods out there we will only introduce the most basic one here. 
A graph convolutional layer \cite{GCNConv} consists of a learnable
weight matrix followed by an aggregation step, formalized by
\begin{equation}
	\mathbf{X}^{\prime} = \mathbf{\hat{D}}^{-1/2} \mathbf{\hat{A}}
	\mathbf{\hat{D}}^{-1/2} \mathbf{X} \mathbf{\Theta}
\end{equation}
where for a given graph $G=(V,E)$, $\hat{A} = A + I$ denotes the
adjacency matrix with added self-loops for each vertex, $D$ is
described by $\hat{D}_{ii} = \sum_{j=0} \hat{A}_{ij}$, a diagonal
matrix displaying the degree of each node, and $\Theta$ denotes the
learnable weight matrix. Added self-loops enforce that each node
representation is directly dependent on its own preceding one. The
number of graph convolutional layers stacked equals the radius of
relevant nodes for each vertex within the graph.

The update rule for each node is given by a message passing scheme
formalized by
\begin{equation}
	\mathbf{x}^{\prime}_i = \mathbf{\Theta} \sum^{N}_{j}
	\frac{1}{\sqrt{\hat{d}_j \hat{d}_i}} \mathbf{x}_j
\end{equation}
where both $\hat{d}_i, \hat{d}_j$ are dependent on the edge weights
$e_{ij}$ of the graph. With simple, single-valued edge weights such as
$e_{ij}=1 \text{ }\forall (i,j)\in E$, all $\hat{d}_i$ reduce to
$d_i$, i.e., the degree of each vertex $i$. We denote this type of
graph convolutional neural layers with \textsc{GCNConv}.

While in this initial formulation of a GCNConv the node-wise update
step is defined by the sum over all neighbouring node representations,
we can alter this formulation to other message passing schemes.  We
can rearrange the order of activation function $\sigma$, aggregation
$\mathrm{AGG}$, and linear neural layer $\mathrm{MLP}$ with this
formulation as proposed by \citet{GENConv2020}:
\begin{equation}
	\mathbf{x}_i^{\prime} = \mathrm{MLP} \left( \mathbf{x}_i +
	\mathrm{AGG} \left( \left\{
	\mathrm{\sigma} \left( \mathbf{x}_j + \mathbf{e_{ji}} \right) +\epsilon
	: j \in \mathcal{N}(i) \right\} \right)
	\right)
\end{equation}
where we only consider
$\sigma \in \{\mathrm{ReLU}, \mathrm{LeakyReLU}\}$. We denote this
generalized layer type as \textsc{GENConv} following the notation of
PyTorch Geometric \cite{PytorchGeometric}.  While the reordering is
mainly import for numerical stability, this alteration also addresses
the vanishing gradient problem for deeper convolutional networks
\cite{GENConv2020}. Additionally, we can also generalize the
aggregation function to allow different weighting functions such as
learnable $\mathrm{SoftMax}$ or $\mathrm{Power}$ for the incoming
signals for each vertex, substituting the averaging step in
\textsc{GCNConv}. Hence, while \textsc{GCNConv} suffers from both
vanishing gradients and signal fading for large scale and highly
connected graphs, each propagation step in \textsc{GENConv} emphasizes
signals with values close to $0$ and $1$. The same convolutional
filter and weight matrix are applied to and learned for all nodes
simultaneously. % , and the resulting information\todo{Which information?
  % Specify} hold no information on their own connectivity.
We further employ another mechanism to avoid redundancy and fading
signals in stacked graph convolutional networks, using residual
connections and a normalization scheme \cite{DeepGCN2019,
  DeeperGCN2020}.  The residual blocks are reusable and can be stacked
multiple times. % The structure of the GNN architecture we use is shown
% in Figure~\ref{fig:ResGraphConvBlocks}.

% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=0.5\columnwidth]{figures/ResGraphConvBlocks.png}}
% 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
% 	\label{fig:ResGraphConvBlocks}
% \end{figure}

% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=1\columnwidth]{figures/full_model_all_layers.png}}
% 	\caption{Residual architecture built by \citet{DeepGCN2019} and \citet{DeeperGCN2020} enabling deeper graph convolutional models}
% 	\label{fig:FullModelAllLayers}
% \end{figure}

\subsubsection{Combined prediction model}
Combining half-twin and graph convolutional neural networks, we map
all protein representations to their respective node features,
initializing the graph convolutional update steps. The resulting
representations are used for a similarity prediction.
% as presented in
% figure ~\ref{fig:HalfTwinNetwork}
When combining ontology and molecular features with or without the
graph model, we concatenate both protein features and both drugs
features, before plugging them into the graph model for the similarity
computation. An overview of the model architecture, combining both
feature types, is shown in Supplementary Figure
\ref{fig:FullModelAllLayers}\todo{Move to supplement}. Here
the original representations are transformed by a dense layer and then
used as input of a stack (with height 3) of residual graph
convolutional blocks.

\subsubsection{Hyperparameter tuning}
As the number of drug-targets are sparse with respect to the number of
both drugs and proteins considered, the training, validation and
testing datasets are imbalanced. As there are only $22,336$ links in
the considered STITCH subset, the ratio
\begin{equation}
	w:= \frac{\#drugs \cdot \#proteins}{\#dti\_links} \approx 360,
	\label{weight}
\end{equation}
consequently needs suitable compensation in the computed loss function and
appropriate metrics for the evaluation.

Therefore, we weight all positive drug-protein pair samples with this
ratio by introducing the following loss function with respect to 
binary cross-entropy:
\begin{equation}
	l(x,y) = - w \left[ y \cdot \log x + (1 - y) \cdot \log (1 - x) \right]
\end{equation}
for a given prediction $x$ and target $y$, and positive weight $w$
defined by equation \eqref{weight}. We average this loss among all
drug-protein pairs in the training set, leading to a stable
environment for the \textit{Adam} optimization algorithm
\citep{Adam2014}. We implemented a 5-fold cross validation split among
the proteins. Furthermore, we used early stopping in the training
process.

To find the best hyperparameter configuration for the proposed model,
we performed a grid search to find the most expressive and
non-redundant representation. We pretrained the bottom-up and the
top-down model separately and aimed at best performing models with
respect to our evaluation metrics. We optimized embedding sizes, depth
of the neural network, optimizer, learning rate and layer types using
an extensive, manual grid search. Starting from shallow feature
transformations with an embedding size of $10$, we scaled the network
up to residual structures with up to $10$ hidden layers leading to
embeddings of size $4000$, testing different network widths and
learning rates for each configuration.

\subsection{Evaluation and metrics}

\subsubsection{Splitting schemes}

As DTI prediction is dependent on both drugs and proteins, there are
multiple ways of determining training, validation and testing sets of
pairs to evaluate each model. For cross-validation, we can perform the
split over DTI pairs, over drugs and over proteins, respectively; when
splitting over drugs or proteins, the entities (drugs or proteins) are
separated and all their associations included in the split.
Only when splitting by protein or drug are unseen entities guaranteed to
be shown to the model in the validation and testing
phase. The models resulting from the different splitting schemes may
have different expressiveness and exploit different information in DTI
prediction, as different information is known in the training and
testing phase.

\subsubsection{Metrics}
\label{sec:evaluation}
To assess each model, we compute a variety of common metrics for
binary classification. As the datasets are highly imbalanced, we use
the area under the receiver operating characteristic curve (AUROC) on
training, validation and testing split.
% We calculated the true positive rate (TPR), false positive rate
% (FPR), and precision score. We will further compute true positives
% (TP), false positives (FP), false negatives (FN) and finally true
% negatives (TN).

We calculate the AUROC by computing true positive rate at various
false positive rate thresholds and use trapezoidal approximations to
estimate the area under the curve. We refer to this measure as
$\textrm{MacroAUC}$.

We also calculate the $\textrm{MicroAUC}$ score. For given lists $D$
and $P$ of drugs and proteins, respectively, and a set of known
interactions $Int := \{(d_i, p_i) \}$, \textit{MicroAUC} is calculated
as the average per entity \textit{AUROC} score. For example, the
protein-centric score can be formalized as: given labels
$l:D\times P \rightarrow \{0,1\}$ and predictions
$y:D\times P \rightarrow [0,1]$, we define
\begin{equation*}
  \textrm{MicroAUC}'_p(l,y) := \underset{p\in P}{mean}\left(\left\{
      \text{AUROC}(\{ (l(d_i, p), y(d_i,p))| d_i\in D\})
    \right\}\right)
\end{equation*}

In some cases, the $MicroAUC$ score may not be defined as in some
datasets some proteins or drugs have no interactions, leading to an
infeasible $TPR=0$ for all thresholds and an undefined $AUROC$ score
for that entity. For those entities, we impute the $MicroAUC$
interpolating linearly, by using the accuracy for this
subset:

\begin{equation*}
	\textrm{MicroAUC}_p(l,y) := 
	\begin{cases}
		MicroAUC'_p(l,y) & \text{if }\sum_{d_i\in
                  D}l(d_i,p)\neq 0\\
		Accuracy(l,y)&otherwise
	\end{cases}
\end{equation*}
Drugs and proteins can be interchanged in this formulation. We refer
to the different measures as protein-centric microAUC ($MicroAUC_p$)
and a drug-centric microAUC ($MicroAUC_d$).

Further, we choose AUROC over area under precision recall curve (AUPRC) as the latter is prone to data imbalance \cite{Jeni2013} and thus sub-sampling biases, undermining cross-approach evaluation.\todo[inline]{More Information? Right place?}

\section{Results}

\subsection{\name{}: computational model to identify drugs that target a
  protein}

% As in machine learning inference is derived from the underlying data,
% models and data are naturally and intrinsically linked. Thus, the more
% we understand about the pitfalls and biases within the data, the more
% we can try to bypass these difficulties. We will hereby abbreviate,
% e.g., a cross validation splitting scheme as \glqq split\grqq{}
% determining the train, validation and test subset of a given dataset.

We developed \name{} as a computational model to predict drug--target
interactions. Specifically, given a protein, \name{} will identify and
rank drugs that likely target this protein. \name{} combines two types
of features: structural information for drugs and proteins that can be
used to determine if the drug and protein physically interact, and
information about phenotypic effects of drugs and changes in protein
function that may ``localize'' on an interaction network (i.e.,
neighboring nodes will share some of these features or are
phenotypically similar). As structural features, \name{} uses
structural representations of drugs from the SMILES transformer
\cite{SmilesTransformer} and representations of protein amino acid
sequences from DeepGOPlus \cite{DeepGoPlus}.  \name{} learns
representations of drug effects and protein functions using the
ontology-based machine learning method DL2Vec \cite{DL2vec2020} and
ontology-based annotations of drugs and proteins.

We construct a graph with proteins as nodes and protein-protein
interactions as edges, mapping the protein features to each target as
node features. \name{} then propagates information among the PPI network
utilizing graph convolutional steps, calculates the similarity of drug
and protein representations, and predicts whether there is an
interaction. The full workflow scheme is depicted in
figure~\ref{fig:ModelWorkflow}

% Note that this approach can be easily generalized, profiting from
% other protein function and phenotype representation methods.

\begin{figure}[!tpb]
	\centering
	\includegraphics[width=0.9\columnwidth]{figures/model_workflow.png}
	\caption{Full DTI prediction model based on the pretrained
          learnable feature transformations (LFT) for either molecular
          structure or ontology based features. The transformed
          protein representations are added to each corresponded
          protein as node features for the graph convolutional
          steps.}
	\label{fig:ModelWorkflow}
\end{figure}

We evaluate our model's ability to identify drug--target interactions
using different approaches and datasets. First, we perform a
cross-validation over proteins and validate our results. A
cross-validation over proteins aims to evaluate how the model performs
when tasked to identify drugs that may target a ``novel'' protein,
i.e., one not seen during training, or a protein for which a drug that
targets it should be predicted.

\begin{table}[ht]
  \centering
  \begin{tabular}{|p{1.1cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|p{0.6cm}|p{0.7cm}|}
    \hline
    &\multicolumn{4}{c|}{(a) STITCH results}& \multicolumn{4}{c|}{(b) Yamanishi results}\\
    \hline
    \name{} results&\multicolumn{4}{c|}{PPI graph}&\multicolumn{4}{c|}{PPI graph}\\
    &\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
    &Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
    \hline
    MolPred&$0.69$&$0.65$&$0.69$&$0.67$&$0.66$&$0.67$&$0.66$&$0.64$\\
    \hline
    OntoPred&$0.88$&$0.87$&$0.92$&$0.93$&$0.80$&$0.79$&$0.83$&$0.82$\\
    \hline
    \name{} & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$& $0.83$ & $0.82$&$\mathbf{0.84}$&$\mathbf{0.84}$\\
    \hline
  \end{tabular}
  \caption{\label{tab:Results}Results for \name{} on the STITCH
    and Yamanishi datasets evaluated with 5-fold cross-validation. We 
    call the model using only molecular features
    \textit{MolPred} and the model using only ontology-based
    features \textit{OntoPred}. \name{} combines both types of
    features.}
\end{table}

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein split;
we then selected the best-performing models (with respect to
$MicroAUC_p$, see Section \ref{sec:evaluation}), and retrain them from scratch in a
5-fold protein-split cross-validation on the Yamanishi benchmark
dataset to avoid validation overfitting and yield more realistic
testing results. To evaluate the influence of the different features
separately, and to determine whether they ``localize'' on the PPI
graph (and therefore can be exploited successfully by the graph neural
networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately. We comparing the molecular (MolPred) and phenotype-based
(OntoPred) prediction model, and a combination of both where we
concatenate both types of features.  Table \ref{tab:Results} shows the
results of these experiments.

We find that the model using ontology-based features
(\textit{OntoPred}) is showing better performance on STITCH compared
to using only molecular features. We also observe that only the model
using ontology-based features results in increased performance when
incorporating the PPI graph. This increase can be observed with
different graph neural network architectures and configurations
(Supplementary Table X\todo{Add to Supplement!}). While the {\em
  GCNConv} and {\em GENConv} architecture already shows some minor
improvement, the use of \textit{ResGraphConv} results in larger
performance improvements.  {\em ResGraphConv} blocks add a large
amount of additional learnable parameters to the network, leading to
more expressive power. To test whether the observed improvement is due
to the number of learnable parameters added or the result of better
exploiting the information about PPIs, we experiment with a graph
model in which all graph convolutional neural layers in the residual
blocks are removed, resulting in a model with similar parameters but
without the ability to use graph-based information. This pruned
network, with no information on the protein--protein interactions,
reached very similar results to the original \textit{OntoPred} model
and showed no improvement (Supplementary Table X\todo{Add!}).

The improvements when including the graph are only provided by the
\textit{GENConv} graph convolution scheme which includes the
{\em ResGraphConv} blocks; \textit{GCNConv} and other graph
convolutional methods fail to achieve any gain in comparison to the
plain \textit{OntoPred} performance even when combined with the
residual blocks. The discrepancy between {\em GENConv} and other graph
convolutional methods may be the result of numerical instability and
fading signals \cite{DeeperGCN2020}.

Our results demonstrate that the inclusion of graph information
can increase performance when ontology-based features are used but not
when molecular features are used alone. This observation allows us to
conclude that information about protein functions localizes on the
graph whereas molecular features do not.

% despite predicting per protein is
% rather counter-intuitive as there only limited drug-targets
% \citep{Overington2006}, and thus novel drugs are more likely to arise
% than novel targets. Yet, we aim to find all interacting drugs for
% existing targets motivating our split choice even further.

\subsection{Protein-centric evaluation}
The goal of \name{} is to find candidate drugs that target a specific
protein; however, so far, we do not evaluate this application but
rather how \name{} would perform in finding plausible drug--target
interactions among all possible interactions (since we use the
MacroAUC as our main evaluation measure). This evaluation does not
correspond to the application of \name{} in finding drugs that target a
specific protein.  To provide a better estimate on how \name{} performs
for individual protein targets, we use micro-averages between proteins
and compute the \textit{MicroAUC} (see Section \ref{sec:evaluation});
to determine {\em MicroAUC}, we average the performance (true and
false positive rates) per protein instead of across all drug--protein
pairs; the resulting measure can therefore better estimate how \name{}
performs when tasked with finding a drug that targets a specific
protein.

% Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively. As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.

% \subsection{Baseline model}
Furthermore, we hypothesize that it may be possible for a machine
learning model to exploit biases in drug--target interaction data to
achieve relatively high prediction performance without obtaining a
biologically meaningful signal.  For example, hub proteins may have a
large number of interactions, or certain drugs interact with many
proteins, and preferentially predicting these interactions may
increase predictive performance even in the absence of any biological
features.  To test this hypothesis, we design a ``na\"ive'' baseline
model that predicts the same list of proteins for each drug based only
on the number of known drug--target interactions for a
protein. Specifically, given lists $D$ and $P$ of drugs and proteins
and a set of known interactions $\mathcal{I} := \{(d_i, p_i) \}$, we
construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$
with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$ predicts all drugs to interact with
the top $k$ targets with respect to the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.

The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for all proteins; consequently,
this na\"ive predictor does not rely on any biological features and
will not predict any novel information about interactions between
drugs and proteins; the na\"ive predictor only exploits imbalances in
the evaluation set to make predictions that may perform well.  The way
in which we formulated the na\"ive predictor, it is not applicable for
a protein split cross-validation as the number of interactions for
each protein in the validation set is unknown.

We apply this naive predictor on both the STITCH and Yamanishi
datasets, using the full datasets as well as a 5-fold cross-validation
over drugs and over drug--protein pairs to compare the prediction
results directly to \name{}.  For each fold, we gradually increase $k$
to determine the best performance for each fold. Using the full
dataset, drug--target split, and drug split, we obtain the following
MacroAUC results: for the STITCH database, we obtain a performance of
$0.76$ on the whole dataset, $0.70$ for the drug--target pairs and
$0.73$ in case of the drug splitting scheme; on the Yamanishi dataset,
we obtain MacroAUC scores of $0.88$, $0.84$ and $0.85$, for the total
dataset, drug--target pair and drug split, respectively.  The na\"ive
predictor shows higher performance on the Yamanishi dataset than on
STITCH, and a substantial gain in comparison to an expected random
predictor on both datasets. In the following, we utilize this nai\"ve
predictor as baseline to compare its performance to state of the art
models and \name{}.

For comparison with the state of the art methods, we chose the best
performing methods for drug--target interaction prediction that were
previously evaluated on the Yamanishi benchmark dataset. These methods
include DTIGEMS+ \cite{DTIGEMS2020} and DTI-CDF \cite{DTI-CDF2019}
which have showed superior results in comparison to numerous
works. Furthermore, we added DTINet \cite{DTINet2017} as method for
comparison which has been used to develop a number of methods such as
NeoDTI \cite{NeoDTI2019} with similar methodology.

We evaluate all models on their recommended splitting scheme choice,
hyperparameters and folds in cross-validation, measuring their
respective AUROC. We further evaluate each model by performing a
protein-wise cross-validation determining the MacroAUC and
$MicroAUC_p$. For this evaluation, we allow
sub-sampling of negatives for the training process
but not for the validation and testing phase as real world
applications of these models would have to deal with possibly
imbalanced data.

\begin{table}[ht]
  \centering
  \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \hline
    Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}\\

    &Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
    \hline
    Naive predictor&Drugs&$0.85$& --&--\\
    DTINet&DP pairs&$0.91$&$0.74$&$0.67$ \\
    DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
    DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
    \name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
    \hline
  \end{tabular}
  \caption{\label{tab:comparison1} Comparison of \name{} with
    state of the art drug--target interaction prediction methods on
    the Yamanishi dataset; we evaluated the original and the
    protein-based split in a cross-validation}.
\end{table}

The results of our experiments are summarized in Table
\ref{tab:comparison1}; we calculated the performance of all compared
methods over their original splitting scheme and over a protein
split. We find that there is a large difference in performance when
evaluating over a drug--target pair split compared to a protein split,
with generally higher performance achieved when using the drug--target
pair split.  Second, when evaluating the same methods over a protein
split, we find a substantial performance difference in comparison to
the splitting scheme used in the original evaluation of each
method. DTI-CDF was originally evaluated on all three splitting
schemes underlining this point \cite{}. While \name{} provides
comparable performance to the na\"ive predictor and DTI-CDF in terms
of MacroAUC, it yields considerably better results with respect to
$\textrm{MicroAUC}_p$. We also find that methods that are trained
using a protein split generally result in higher $\textrm{MicroAUC}_p$
than methods trained using a drug--target pair split, indicating that
they may generalize better to unseen protein targets whereas methods
trained on a drug--target split potentially exploit hidden biases and
therefore generalize less well.

\begin{table}[ht]
  \centering
  \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
    \hline
    Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}\\

    &Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
    \hline
    Naive predictor&DP pairs&$0.79$&-- &--\\
    DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
    DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
    DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
    MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
    \name&Proteins&0.85 & $\mathbf{0.85}$& $\mathbf{0.82}$\\
    \hline
  \end{tabular}
  \caption{\label{tab:comparison2} Comparison of \name{} with
    state of the art drug--target interaction prediction methods
    on the BioSnap dataset; we evaluated the original and the
    protein-based split in a cross-validation}.
\end{table}

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluation on other datasets.  Following the results of MolTrans
\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
MolTrans itself on the BioSnap dataset \cite{} and compared it to our
``na\"ive'' predictor as well as \name{} (see Table
\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
pair and the protein split; we were able to reproduce the MolTrans
results (Table \ref{tab:comparison2}), showing a substantial
difference based on the splitting scheme. We additionally computed the
$MicroAUC_p$ score for all considered methods, leading to similar 
results as observed on the Yaminishi dataset.

In all our experiments, \name{} has the highest Micro $AUC_p$,
demonstrating that \name{} can improve over other methods in the task
of identifying drugs that target a specific protein. Several methods
achieve a higher Macro AUC than \name{}, in particular when evaluated
using a drug--target pair splitting scheme; our results with the
``na\"ive'' prediction method show that it may be possible that models
trained on a drug--target split utilize certain biases in the dataset
without necessarily producing novel biological insights.

\section{Discussion}

\subsection{``Bottom-up'' and ``top-down'' prediction of drug-target interactions}

There are many computational methods to predict drug--target
interactions. They can broadly be grouped in two types; the first,
which we refer to as ``bottom-up'' approaches, start from molecular
information about a drug and protein and predict an interaction based
on their molecular properties; the second, which we refer to as
``top-down'' approaches, start from observable characteristics of an
organism and infer drug--target interactions as the putative molecular
mechanisms that explain these observations.

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. Molecular information can be used
to directly determine whether two molecules (such as a drug and
protein) have the ability to interact, whereas information about
phenotypic consequences of a drug (drug effects) or disruption of a
protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular information will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Our results (Table
\ref{tab:Results}) confirm this hypothesis and demonstrate that
molecular features do not benefit from including the interaction
network whereas the indirect, top-down features benefit from the
propagation over the interaction network.

% \todo[inline]{Methods:}
% Through the very nature of the graph convolutional neural network, we
% build the transformed representation for all proteins in every
% forwarding step of the model. Note particularly, that the same
% convolutional filter and weight matrix are applied to and learned for
% all nodes simultaneously. By construction, for a single drug we can
% compute and predict all its interactors in a single run of the model,
% leading to significantly less computing time.

There are other types of indirect features that could be added to our
model. A common feature that may be added are drug indications which are
predictive of drug--target interactions \cite{}; however, we do
not include them in our model as including drug indications would
allow our model to make many trivial predictions based only on
remembering which targets are often used for which indication;
including network information would likely benefit predictions based
on drug indications because different drugs may target the same
pathway through different mechanisms.

Combining bottom-up and top-down approaches in a single model can
follow different strategies. Interaction networks are used widely to
determine indirect effects of molecular changes \cite{} and predict
drug--target interactions \cite{}. Our work relies on graph neural
networks as a way to combine qualitative information about
interactions with additional features (molecular interaction,
phenotypic and functional features); even if only some of these
features benefit from the information the graph provides, graph neural
networks will allow further extension of our model with additional
features in the future.

% Graph neural networks \cite{} have the ability to
% represent interaction networks

% we built protein function and ontology based features based on DL2vec

% Ontology derived protein function focused features are highly predictive for dtis

% We built a versatile template for various features to test localization on the PPI graph

% normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability

% molecular features work for all (most?) drugs, top-down features only
% for those that have a side-effect profile

% \begin{itemize}
% 	\item \name{} comparison to other approaches on BioSnap prone to bias, as $ BioSnap: drugs 
% 	\times prots = 5500 \times 3000 $, while only 
% 	$\approx$ 1000 drugs have Sider Annotation, and 
% \end{itemize}

\subsection{The challenge of evaluating drug--target interaction
  predictions}

One major component of our experiments was to determine how the
information that is available to a machine learning model during
training affects the performance of the model. Similarly to previous
work \cite{}, we find significant differences in predictive
performance across different splitting schemes.

The most common scheme for drug--target interaction prediction is the
split over drug-target pairs \cite{Survey2018} where it may happen
that most drugs and targets that are including in the model's
validation and testing phase have also been included in the training
phase (as part of other drug--target pairs). This scheme is prone to a
number of biases. If the number of interactions for a drug or protein
are imbalanced, i.e., some drugs or proteins have many more
interactions than others, these will be seen more often during
training and they will likely also have more interactions in the
testing and validation sets; because some entities have more
interactions, i.e., they are more likely to interact, any model that
preferentially predicts these as interaction partners will improve its
predictive performance. While this accurately captures the
distribution, predicting based on biases in the number of interaction
partners is not desirable when applying the model to novel
entities. We have demonstrated that even a ``na\"ive'' classifier that
makes predictions only by exploiting the imbalanced number of
interaction partners can achieve performance close to state of the art
methods (when measuring Macro AUC). When training a machine learning
model on such imbalanced data, it will eventually overfit to this
imbalance.  Splitting by entity (protein or drug) can reduce the
impact of this spurious correlation but not reduce it entirely,
because similar entities will still exhibit similar interaction
patterns. In our experiments, we observed the impact of splitting
training and evaluation sets by protein as a decrease in overall
performance (Macro AUC), providing some evidence that models trained
using this splitting scheme are less sensitive to overfitting to this
type of bias.

The way in which training and evaluation data is generated is related
to how the model is evaluated. An evaluation based on Macro AUC
evaluated the application scenario where a set of drugs and proteins
are given, and out of all possible pairs, the more likely interactions
need to be identified. However, this does not correspond to most
scenarios for drug repurposing where a drug that targets a {\em
  specific} protein (e.g., a protein involved in a disease) needs to
be identified. We use an evaluation measure based on micro-averages
per protein (Micro AUC$_p$) to evaluate this scenario, and we often
find substantial differences in predictive performance when evaluating
with Macro AUC and Micro AUC$_p$; generally, models that are trained
using a split over drug--target pairs perform worse in Micro AUC$_p$
than models that use a protein-based split, further providing evidence
that a drug--target split results in overfitting to dataset biases.

Finally, a potential source of differences in model performance is how
negatives are identified and treated during evaluation (and
training). There are few large sets of validated negative drug--target
interactions; consequently, many models (including \name{}) use all
unknown interactions as negatives. As there are many more negative
than positive interactions, negatives are then sub-sampled during
training resulting in a training set that is balanced between
positives and negatives (or a certain ratio is preserved). While this
is a reasonable strategy to deal with imbalanced data, it may be
problematic when the same sub-sampling is applied on the model's
evaluation set because it over-simplifies the evaluation process. The
performance differences is not usually visible when using ROC curves
but results in unrealistically high precision and therefore high area
under a precision-recall curve.

In summary, drug--target interaction prediction is not a single
computational problem in bioinformatics but a set of related
problems. Let $P$ be a set of proteins $P$ and $D$ a set of drugs; one
task can be to identify arbitrary pairs $(p,d)$ with $p \in P$ and
$d \in D$ that interact, another to identify a set of interacting
drugs for each $p \in P$, and yet another to identify a set of
proteins for each drug $d \in D$. The first task may be useful when no
particular drug or protein is considered; the second task when
searching for a drug that targets a specific (disease-associated)
protein; and the third when aiming to find new applications for a
given drug. The first task would best be evaluated using a Macro AUC,
the second and third using a Micro AUC$_p$ and Micro AUC$_d$.

% The second most prevalent arrangement is the split
% over drugs, while only close to none is aiming on a protein split.
% However, the first and second splitting scheme are exposed to the
% first dataset bias and are hence more likely vulnerable to
% transductive inference by just predicting recently seen structures,
% rather than implementing inductive inference and generalizing over the
% drug representations. Second, these two strategies are more
% susceptible to the second bias, as only in these cases the model may
% overfit on the number of existing interactions for a single protein,
% while in the third scheme the number of interactions of the test
% proteins is entirely unknown during training process.

% We designed a
% na\"ive predictor and find that it can achieve a performance only
% slightly below state of the art drug--target prediction methods. We
% investigated this finding further and

% In the presence of imbalanced

% Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it

% Running split over proteins is harder than, drug and drug protein pair split (see below table)

% this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)

% Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

% microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug

% \todo[inline]{Likely discussion:}
% The aim of \name is to predict candidate drugs that target a given
% protein; the challenge is to develop a training and evaluation scheme
% that does not simply overfit to the inherent biases in training and
% testing data.
% In general, when performing cross-validation for DTI prediction, the
% options are to split over 
% \begin{enumerate}
% 	\item split over drugs,
% 	\item split over drug--target pairs, or
% 	\item split over proteins
% \end{enumerate}
% where the first and third option concern splitting drugs and proteins,
% respectively, into train, validation and test sets, and arranging the
% corresponding drug-target interactions. They ensure that at least
% parts of the interactions are not seen during training and evaluate
% either how well targets are predicted for unseen drugs or unseen
% proteins. Hereby, different training and prediction schemes lead to
% divergent expressiveness of the resulting model.

% A naive predictor (ranking proteins) and predict each drug similarly
% achieves cutting edge performance ($87.5$ AUROC for whole dataset,
% $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
% prot focused microAUC possible. $\rightarrow$ hub proteins

% Yamanishi Dataset is only partially suitable for comparing results, if
% everybody just derives a suitable subset (DTIGEMS)

% This also applies to drug target affinity prediction. We were hereby
% able to roughly reproduce the results from MolTrans (Bioinformatics)
% on BioSnap

% %\subsection{Features to use in predicting drug targets}

%\todo[inline]{Discussion:}
% The lack of stratification, only impacts the not
% considered area under precision recall curve (AUPRC) and not the Macro
% AUROC score, while also supporting the expressiveness of $MicroAUC_p$
% with more data points.

%\subsection{\name identifies drugs that target a protein}

% \todo[inline]{Likely discussion:}
% Assuming a hypothetical, perfectly generalizing model built upon an
% unbiased dataset, this very model would yield similar performances for
% all three , not overfitting on the known
% structures. On the other hand, for a hypothetical entirely overfitting
% model, trained on a highly biased dataset, this model would show
% substantial deviations from the original performance over another
% split.

% \todo[inline]{Likely discussion:}
% We emphasize, that all real-world models are prone to some sort of
% overfitting, and unknown, deviant entities in both validation and
% testing set will likely lead to some sort of performance gap for
% relevant metrics. However, a large disparity may hint the biases
% stated above.

\section{Conclusions}

We developed \name{} as a machine learning model that combines
molecular features and functional information with an interaction
network using graph neural networks to predict drugs that may target
specific proteins. In this task, \name{} improves over several state
of the art methods. We demonstrated that functional and phenotypic
information localizes on the interaction network whereas molecular
information does not. Moreover, we showed that drug--target
interaction prediction datasets have some inherent biases that affect
the performance of models. This led us to conclude that drug--target
interaction prediction is not a single computational problem but a set
of multiple problems. Experimental evaluation of drug--target
interaction prediction methods must be carefully designed to reflect
the problem the model aims to solve, and the interpretation of
performance results should be aligned with the specific problem.

% \subsection{Deification of our method}
% \begin{itemize}
% 	\item all AUROC in \% AUROC score on STITCH

% 	\begin{tabular}{|l|r|r|}
% 		\hline
% 		DNNT model&Without graph model&With graph model\\
% 		\hline
% 		MolPred&$69$&$69$\\
% 		PhenomeNETPred&$88$&$92$\\
% 		\hline
% 		MolPred + PhenomeNETPred & $89$ & $93$\\
% 		\hline
% 	\end{tabular}
% 	\item microAUC for MolPred + PhenomeNETPred on graph is about $93+-$
% 	\item and on yamanishi dataset

% 	\begin{tabular}{|l|r|r|}
% 		\hline
% 		DNNT model&Without graph model&With graph model\\
% 		\hline
% 		PhenomeNETPred&$83$&$84$\\
% 		\hline
% 		MolPred + PhenomeNETPred & $83$ & $84.5$\\
% 		\hline
% 	\end{tabular}
% 	\item MicroAUC is about $83$
% \end{itemize}

% \subsection{How to insult other methods}
% \begin{itemize}
% 	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
% 	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug

% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
% 		\hline
% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
% 		\hline
% 		DTINet&DP pairs&$91$&$84.1$&$67.2$\\
% 		DTIGEMS+&DP pairs&$93$& $72.2$& $67.8$ \\
% 		DTI-CDF&Proteins&$83$&$83$&$79$\\
% 		\hline
% 	\end{tabular}
% 	\item A naive predictor (ranking proteins) and predict each drug similarly achieves cutting edge performance ($87.5$ AUROC for whole dataset, $85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No prot focused microAUC possible. $\rightarrow$ hub proteins
% 	\item Yamanishi Dataset is only partially suitable \todo{for
%             comparing results}, if everybody just derives a suitable subset (DTIGEMS)
% 	\item This also applies to drug target affinity prediction. We were hereby able to roughly reproduce the results from MolTrans (Bioinformatics) on BioSnap

% \end{itemize}

% \begin{figure}[!tpb]%figure1
% 	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|r|}
% 		\hline
% 		Approach&Splitting scheme&Original AUROC score&Protein split AUROC&MicroAUC\\
% 		\hline
% 		DeepDTI&Drugs&$87.6$&$75.9$&$70.1$\\
% 		DeepDTA&DP pairs&87.6&$76.7$&$69.4$\\
% 		DeepConv-DTI&DP pairs&88.3&$76.6$&$73.0$\\
% 		MolTrans&DP pairs&89.5&$77.0$&$74.0$\\

% 		\hline
% 	\end{tabular}
% \end{figure}

% \subsection{Tested hypotheses}

% In this work we are testing the following hypotheses:
% \begin{enumerate}
% 	\item Can we build a model that outperforms state of the art approaches, combining top-down and bottom-up approaches?
% 	\item Are interaction networks sufficient to improve the performance of simple molecular predictors?
% \end{enumerate}
% We will test the first hypothesis by building a model that takes both top-down and bottom-up features into account. Thus, we propose a novel approach to combine those mutual exclusive attempts, through the usage of interaction networks, similarity and molecular features. Additionally, we test the latter by building a simple molecular DTI predictor and enhance it under usage of the interaction networks.\\

% \section{Methods}

% \subsection{Models}
% The used model consists of two separate models, that help to fuse together the two methods:
% \begin{enumerate}
% 	\item The molecular predictor
% 	\item The interaction network based predictor
% \end{enumerate}
% We build the molecular predictor by using pretrained, molecular fingerprints models for both drugs and proteins. Regarding proteins, we used the pretrained feature generator from \textit{DeepGoPlus} (\citep{DeepGoPlus}) that was originally designed for protein function prediction and is regarded as state of the art for this purpose. For drugs we used a pretrained fingerprint model from \textit{SMILES transformer} (\cite{SmilesTransformer}), that provides a simple and fast method to compute fingerprints through autoencoder models. The encodings from these two models were funneled into a simple deep neural network (see Figure~\ref{fig:01}) with few fully connected. \\
% The results of that prediction flow into the annotation of the protein-protein interaction (PPI) graph as depicted in (IMAGE). Hereby, the predictions of the molecular predictor are used as node features for the graph, with respect to the given drug. Thus, given a compound-target pair, the nodes of the PPI graph now hold bottom-up features, which can now be processed by the graph learning algorithms. \\
% The PPI graph is processed by different graph convolutional layers, that may underline the importance of either patterns or regions within the graph, to obtain a feature vector for the wanted node. In contrast to learning over whole graphs we perform node classification within the graph. These layers are either graph convolutional layers, that learn a certain kernel over the graph, or attention based. Different layers of both and other types such as were tested.  \\
% The drug-drug interaction features are retrieved by choosing the corresponding row in the adjacency matrix of the graph, thus leading to quite simple features. \\
% For the semantic similarity feature, that once again represents a top-down attribute, we artificially link each drug to its corresponding side effects in the MedDRA hierarchy. Concerning this hierarchy, drug-drug similarity is computed by the Resnik similarity (\cite{Resnik1995}). For the given compound we take the corresponding row of this symmetric similarity matrix. \\

% Thereby, we concatenate these three features together and funnel them into another deep neural network as depicted in figure \ref{fig:02}. This network finally yields our prediction. We hereby perform splits over both drugs and proteins, in order to test and show the discrepancy and increasing difficulty.\\ Implementation was done in PyTorch (\cite{Pytorch}) and is available on Github under \href{github.com/thinnerichs/KAUST-dti-metabol}{github.com/thinnerichs/KAUST-dti-metabol}. Graph learning methods were build with help of PyTorch-Geometric (\cite{PytorchGeometric}), a geometric deep learning extension library for PyTorch, that recently got a lot of attention in the machine learning community. This library gives the potential to use many state of the art graph learning mechanisms, such as plain but effective graph convolution (\citet{GCNConv}), Chebychev kernels (\cite{ChebConv}), ARMA kernels (\cite{ARMAConv}), translation-invariant operators (\cite{FeaStConv}), attention mechanisms (\cite{GATConv}), random walks (\cite{APPNPConv}) and mixtures of the latter two (\cite{SAGEConv}). The performance of these various layer types were tested for this particular problem, as discussed in the results section.

% \begin{figure}[!tpb]%figure1
% 	\centerline{\includegraphics[width=0.4\textwidth]{figures/MolecularPredictor.png}}
% 	\caption{Molecular predictor based on the generated features from DeepGoPlus and SMILES transformer.}\label{fig:MolPred}
% \end{figure}

% \begin{figure}[!tpb]%figure2
% \centerline{\includegraphics[width=0.4\textwidth]{figures/InteractionNetwork.png}}
% \caption{Deep neural network that predicts based on drug-drug interaction features and semantic similarity features over side effects for drugs, and graph convolution over protein-protein interaction networks for proteins. Protein and drug features are represented by blue and red, respectively.}\label{fig:02}
% \end{figure}

%\section{Discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
%     as it may ignore the figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgements}
We acknowledge the use of computational resources from the 
KAUST Supercomputing Core Laboratory.

\section*{Funding}

This work was supported by funding from King Abdullah University of
Science and Technology (KAUST) Office of Sponsored Research (OSR)
under Award No. URF/1/3790-01-01 and URF/1/4355-01-01.

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
\bibliography{citations}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: