### 2021-04-21 paper/cover_letter/cover_letter.tex
Update download scripts

\documentclass[]{letter}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{TO\_ADDRESS}
\address{ADDRESS}

Dear Editor,

Please find our manuscript entitled "DTI-Voodoo: machine learning over interaction networks and
ontology-based background knowledge predicts drug--target
interactions" for your consideration. We therefore developed a novel method for drug--target interaction prediction (DTI) utilizing both ontology-encoded phenotypical and molecular features for both drugs and proteins. We further combine these methods with interaction networks, aiming for a more protein-centric prediction and evaluation.

We designed DTI-Voodoo under usage of graph convolutional neural network, exploiting the ``guilt-by-association`` principle and incorporate both top-down and bottom-up features into protein-protein interaction networks. Furthermore, we show that common DTI datasets share immanent biases, majorly influencing predictive DTI model performance evaluations. Thus, we propose a modified evaluation scheme, showing that DTI-Voodoo advances significantly over other state of the art DTI prediction approaches over several datasets.

The importance of DTI prediction for other challenges such as drug repurposing and repositioning makes DTI-Voodoo highly influential for a variety of researchers. Eventually, DTI-Voodoo's versatile methodology may be used over alternative protein-based representations to improve performances within other predictive tasks in the realm of bioinformatics and computational biology.

Therefore, we believe that
DTI-Voodoo will be well suited for the broad readership of
Bioinformatics.

Please do not hesitate to contact us should you require any further
information.

With best wishes,
Tilman Hinnerichs and Robert Hoehndorf.\\

\signature{}

%enclosure listing
%\encl{}

\end{letter}
\end{document}