### 2021-04-27 paper/cover_letter/cover_letter.tex
.

\documentclass[]{letter}

\begin{document}
% If you want headings on subsequent pages,
% remove the ``%'' on the next line:
% \pagestyle{headings}

\begin{letter}{TO\_ADDRESS}
\address{ADDRESS}

Dear Editor,

Please find our manuscript entitled "DTI-Voodoo: machine learning over
interaction networks and ontology-based background knowledge predicts
drug--target interactions" for your consideration.

DTI-Voodoo is a novel method for drug--target interaction (DTI)
prediction utilizing both ontology-based information about phenotypic
effects of drugs (side effects), and molecular features that encode
for interactions between drugs and proteins. We compare DTI-Voodoo to
seven state of the art DTI prediction methods on two datasets, and
significantly improve over all of them in our evaluation.

Moreover, we introduce several novel aspects to DTI prediction: First,
we identify, analyze, and correct for a number of inherent biases in
drug--target interaction prediction datasets; for this purpose, we
propose a ``naive'' classifier which does not make predictions based
on any biological information but achieves performance close to state
of the art methods.  Second, DTI-Voodoo combines indirect information
such as side effects and molecular properties of both drugs and
proteins in a novel way using an interaction network. Using this
network, we can demonstrate that side effect information ``localizes''
on the interaction network, i.e., that side effects are enriched on
modules consisting of interacting proteins, whereas molecular features
do not benefit from information the graph.

We believe that our results, as well as the novel methods we
introduced in developing DTI-Voodoo, makes our manuscript highly
suitable for the readers of Bioinformatics.

Please do not hesitate to contact us should you require any further
information.

With best wishes,
Tilman Hinnerichs and Robert Hoehndorf.\\

\signature{}

%enclosure listing
%\encl{}

\end{letter}
\end{document}